---
Author:sai sarvani 
output:
  pdf_document: default
  html_document: default
---
Introduction
========================================================

In this project, we will be considering red vinho verde wine samples, a unique product from the Minho (northwest) region of Portugal. In particular, the red wine quality is of interest, and this project will focus on exploratory data analysis (EDA) to explore the relationships in one variable to multiple variables.

Moving on, more details on the wine data will be given. Then, exploratory data analysis (EDA) on the data will be performed.

Dataset 
========================================================

As mentioned before, the data is about red wine samples (vinho verde) from Portugal. The data was collected from May 2004 to February 2007 using only protected designation of origin samples that were tested at the official certification entity.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
knitr::opts_chunk$set(fig.width=9,fig.height=5,fig.path='Figs/',
                      fig.align='center',tidy=TRUE,
                      echo=FALSE,warning=FALSE,message=FALSE)

library(ggplot2)
library(GGally) 
library(purrr) 
library(tidyr) 
library(reshape) 
library(knitr)
library(dplyr)
library(corrplot)
```

```{r ECHO=FALSE, Load_the_Data}
# Load the Data
Wine <- read.csv("C:/Users/user/Desktop/R project/wineQualityReds.csv")
Wine<-Wine[,-1]
# Save the original data before making any changes
Wine.orig = Wine
# Change quality to categorical
#Wine$quality = as.ordered(Wine$quality)
## Split wine quality into good wine or bad wine 
#WineBinaryQuality = Wine.orig
#WineBinaryQuality$quality = as.factor(ifelse(Wine.orig$quality>5.5,"Good","Bad"))
#glimpse(Wine)

```


The data contains 1599 observations (wine samples) and 12 attributes or variables related to the wine. The 12 attributes and a description of each variable (attribute) are provided in the table below:

| Variable             | Description                                                                                 |
|----------------------|---------------------------------------------------------------------------------------------|
| Fixed acidity        | Amount of tartaric acid (in grams) per decimeter cubed of wine (dm^3) [g/dm^3]              |
| Volatile acidity     | Amount of acetic acid (in grams) per decimeter cubed of wine (dm^3) [g/dm^3]                |
| Citric acid          | Amount of citric acid (in grams) per decimeter cubed of wine (dm^3) [g/dm^3]                |
| Residual sugar       | Amount of residual sugar (in grams) per decimeter cubed of wine (dm^3) [g/dm^3]             |
| Chlorides            | Amount of sodium chloride (in grams) per decimeter cubed of wine (dm^3) [g/dm^3]            |
| Free sulfur dioxide  | Amount of free sulfur dioxide (in milligrams) per decimeter cubed of wine (dm^3) [mg/dm^3]  |
| Total sulfur dioxide | Amount of total sulfur dioxide (in milligrams) per decimeter cubed of wine (dm^3) [mg/dm^3] |
| Density              | Density of wine [g/cm^3]                                                                    |
| pH                   | Acidity/Alkalinity of the wine                                                              |
| Sulphates            | Amount of potassium sulphate (in grams) per decimeter cubed of wine (dm^3) [g/dm^3]         |
| Alcohol              | Amount of Alcohol by percent volume                                                         |
| Quality              | A score between 0 and 10 based on Sensory Data                                              |


# Univariate Plots Section

```{r ECHO=FALSE, Univariate_Plots}
Wine %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value,fill=key)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins=sqrt(nrow(Wine))) +
  theme(legend.position="none")
```

From Figure 1, one can see that density and pH seem to be symmetric with no heavy tails. This points to the distribution visually looking like a normal distribution. Volatile acidity also shows a somewhat normal distribution. Looking at the remaining variables, one can also see that a number of distributions are skewed (with a right tail), like chlorides, fixed acidity, residual sugar, sulphates among others. This means these distributions have mostly lower end values with a few wines with relatively higher values. For example, looking at residual sugar, one can see a majority of the values fall between 0 and 4 grams of residual sugar per decimeter cubed of wine; however, there are a few values that are above 4 and even some that approach 16.

If a skewed distribution is undesirable, one method is to log transform the variable. We can show this effect by making the x axis on a log scale. The same plots are given below, but with the x axis on a log scale.


    
```{r ECHO=FALSE, Univariate_Plot}
Wine %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value,fill=key)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins=sqrt(nrow(Wine))) +
    theme(legend.position="none") +
    scale_x_continuous(trans='log10')
```

```{r Echo=False, wine_quality}
Wine$oquality = as.ordered(Wine$quality)
ggplot(Wine, aes(x=quality, fill = oquality)) +
  geom_bar(stat="count") +
  geom_text(position = "stack", stat='count',aes(label=..count..), vjust = -0.5)
+
  labs(y="Number of Observations", x="Wine Quality") +
  theme(legend.position="none")
```

From Figure 3, one can see quality is not balanced across its entire range of 0-10. Most of the numbers are around 5 or 6. In other words, there are much more normal wines than very excellent or poor ones. This may make it harder to determine what makes an excellent or poor wine. One could just split the wines into a good vs bad wine quality to help alleviate this issue.

With 5 and 6 having a relatively even number of wines and having the majority of the wines overall, it would make logical sense to split the wines into bad and good wines by 5 or below and 6 or above if one wants to make a binary wine quality variable that is somewhat evenly distributed. This will be explored at later. For now, let us continue with the wine qualities as they are.


# Univariate Analysis

```{r Echo=False}
summary(Wine)
```

The features that characterize the red wine are related to acidity (fixed.acidity, volatile.acidity, citric.acid, pH), sugar (residual.sugar), content of sulfur dioxide which is the substance that (free.sulfur.dioxide, total.sulfur.dioxide).

Considering what we have inferred in the previous section, we have created a new variable that separates good wines from bad ones. The following histogram shows the count of this variable

```{r Echo=False }
Wine$good.wine<-ifelse(Wine$quality>5.5,1,0)

ggplot(Wine,aes(x=good.wine,fill=factor(good.wine)))+geom_bar(stat = "count",
                                                            position = "dodge")+
  scale_x_continuous(breaks = seq(0,1,1))+
  ggtitle("Distribution of Good/Bad Red Wines")+
  theme_classic()
```



# Bivariate Plots Section

```{r Echo=False, Bivariate_Plots}
Wine.orig %>% cor() %>% corrplot.mixed(upper = "ellipse", tl.cex=.8, 
                                       tl.pos = 'lt', number.cex = .8)
```


The following dimmensions are relatively highly correlated:

**total.sulfur.dioxide** with **free.sulfur.dioxide**;
**fixed.acidity** with density and **citric.acid**;

The following dimmensions are relatively correlated:

**alcohol** with **quality** (this might be a candidate for drop, since might be a leak);

The following dimmensions are relativelly highly inverse correlated:

**fixed.acidity** with **pH**;

The following dimmensions are relatively inverse correlated:

**citric.acid** with **pH** and volatile.acidity;


We're interested in wine's quality, so we care about the final 2 columns/rows in order to know which among the variables has the strongest relationship with wine quality. As the heatmap suggests, alcohol has the strongest correlation with wine quality.

Let us see how each variable is related to wine quality. A method to visualize the relationships between variables is with the pairs plot as shown below

```{r}
ggplot(Wine,aes(x=fixed.acidity,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(fixed.acidity[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(fixed.acidity[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(4,16,1))+
  xlab(label = "Fixed Acidity Level")+
  ggtitle("Distribution of Fixed Acidity Levels")+
  theme_classic()
```

```{r}
#Volatile Acidity and Wine Quality
ggplot(Wine,aes(x=volatile.acidity,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(volatile.acidity[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(volatile.acidity[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,1.6,0.1))+
  xlab(label = "Volatile Acidity Level")+
  ggtitle("Distribution of Volatile Acidity Levels")+
  theme_classic()
```



```{r}
#Citric Acid and Wine Quality
ggplot(Wine,aes(x=citric.acid,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(citric.acid[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(citric.acid[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,1,0.1))+
  xlab(label = "Citric Acid Level")+
  ggtitle("Distribution of Citric Acid Levels")+
  theme_classic()
```



```{r}
#Residual Sugar and Wine Quality
ggplot(Wine,aes(x=residual.sugar,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(residual.sugar[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(residual.sugar[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0.5,15.5,1))+
  xlab(label = "Residual Sugar Level")+
  ggtitle("Distribution of Residual Sugar Levels")+
  theme_classic()
```



```{r}
#Chlorides and Wine Quality
ggplot(Wine,aes(x=chlorides,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(chlorides[good.wine==0],na.rm=T)),color="red",
             linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(chlorides[good.wine==1],na.rm=T)),color="blue",
             linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0.01,0.62,0.1))+
  xlab(label = "Chlorides Level")+
  ggtitle("Distribution of Chlorides Levels")+
  theme_classic()
```


```{r}
#Free Sulfur Dioxide and Wine Quality
ggplot(Wine,aes(x=free.sulfur.dioxide,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(free.sulfur.dioxide[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(free.sulfur.dioxide[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,72,8))+
  xlab(label = "Free Sulfur Dioxide Level")+
  ggtitle("Distribution of Free Sulfur Dioxide Levels")+
  theme_classic()
```


```{r}
#Total Sulfur Dioxide and Wine Quality
ggplot(Wine,aes(x=total.sulfur.dioxide,fill=factor(good.wine)))+
  geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(total.sulfur.dioxide[good.wine==0],na.rm=T)),
             color="red",linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(total.sulfur.dioxide[good.wine==1],na.rm=T)),
             color="blue",linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,300,20))+
  xlab(label = "Total Sulfur Dioxide Level")+
  ggtitle("Distribution of Total Sulfur Dioxide Levels")+
  theme_classic()
```


```{r}
#Density and Wine Quality
ggplot(Wine,aes(x=density,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(density[good.wine==0],na.rm=T)),color="red",
             linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(density[good.wine==1],na.rm=T)),color="blue",
             linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0.9,1.1,0.05))+
  xlab(label = "Red Wine Density Level")+
  ggtitle("Distribution of Red Wine Density Levels")+
  theme_classic()
```


```{r}
#PH and Wine Quality
ggplot(Wine,aes(x=pH,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(pH[good.wine==0],na.rm=T)),color="red",
             linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(pH[good.wine==1],na.rm=T)),color="blue",
             linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(2.5,5,0.5))+
  xlab(label = "Red Wine PH Level")+
  ggtitle("Distribution of Red Wine PH Levels")+
  theme_classic()
```


```{r}
#Sulphates and Wine Quality
ggplot(Wine,aes(x=sulphates,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(sulphates[good.wine==0],na.rm=T)),color="red",
             linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(sulphates[good.wine==1],na.rm=T)),color="blue",
             linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(0,2,0.25))+
  xlab(label = "Sulphates Level")+
  ggtitle("Distribution of Sulphates Levels")+
  theme_classic()
```


```{r}
#Alcohol and Wine Quality
ggplot(Wine,aes(x=alcohol,fill=factor(good.wine)))+geom_density(alpha=0.25)+
  geom_vline(aes(xintercept=mean(alcohol[good.wine==0],na.rm=T)),color="red",
             linetype="dashed",lwd=1)+
  geom_vline(aes(xintercept=mean(alcohol[good.wine==1],na.rm=T)),color="blue",
             linetype="dashed",lwd=1)+
  scale_x_continuous(breaks = seq(8,15,1))+
  xlab(label = "Alcohol Level")+
  ggtitle("Distribution of Alcohol Levels")+
  theme_classic()
```

# Bivariate Analysis

from the previous pair plots, we can notice that residual_sugar, free_sulfur_dioxide, total_sulfur_dioxide, and chlorides do not have significant different across difference quality. Thus, volatile_acidity, sulphates, and alcohol have more significate difference across different qualities. The rest of the variables are showing some difference; however, the difference among different qualities is not significant enough to determine all the quality directly by using one variable.

#Lets explore some more bivariate plots with respect to quality of red wine 

# Bivariate Plots Section

```{r echo=FALSE}
make_box_with_dots <- function(variable){
  ggplot(data = wines, aes_q(x = ~quality, y = as.name(variable)))+
    geom_boxplot()+
    geom_jitter(width = 0.7, alpha = 1/10)+
    geom_line(aes(group = 1),
                  stat = "summary",
                  fun.y = median,
                  color = "#E74C3C",
                  size = 1,
                  alpha = 0.8)
}

cond_median <- function(variable){
  print(paste("Median of", variable, "by quality:"))
  print(by(wines[[variable]], wines$quality, median))
}

boxplot_and_median <- function(variable){
  cond_median(variable)
  make_box_with_dots(variable)
}
```

## Fixed Acidity vs. Quality


```{r echo=FALSE}
boxplot_and_median("fixed.acidity")

```

We see a very slight upwards trend of higher quality with higher fixed acidity. However, the extreme quality classes (3 and 8) have less observations than the middle ones, which may make the median value not so accurate. And we see a drop of acidity from 7 to the 8 quality class. Additionally, we see a big dispersion of acidity values across each quality scale. This may be a indicator that the quality cannot be predicted based only on the value of acidity and is the result of a combination of more variables.

## Volatile Acidity vs. Quality

```{r echo=FALSE}
boxplot_and_median("volatile.acidity")

```

Having in mind the same limitations as referred for the Fixed Acidity (extreme classes with less observations and variability inside the same quality class), we can see a more obvious trend. Lower volatile acidity seems to mean higher wine quality.

## Citric Acid vs. Quality

```{r echo=FALSE}
boxplot_and_median("citric.acid")

```

Higher citric acid seems to mean a higher quality wine. The citric acid is always in low concentrations and in the univariate plots we saw that the distribution peaked at the zero value.

Let's see which proportion of wines has zero citric acid. For all the wines that proportion is:

```{r echo=FALSE}
# How many have zero citric acid
# Total Percentage
sum(wines$citric.acid == 0) / nrow(wines)
```

For each quality class the proportions are:

```{r echo=FALSE}
# Grouped Percentage
group_by(wines, quality) %>%
  summarise(n_zero = sum(citric.acid == 0)/n())
```

We see a decreasing proportion of wines with zero citric acid on the higher quality classes.

So, this reinforces the first impression that the higher citric acid concentration relates to higher quality wines.

## Residual Sugar vs. Quality

```{r echo=FALSE, warning=FALSE}
boxplot_and_median("residual.sugar")

#removing the top 10% to be able to have a better look
boxplot_and_median("residual.sugar")+
  ylim(NA, quantile(wines$residual.sugar, 0.90))

```

Residual sugar seems to have a low impact in the quality of the wine.

## Chlorides vs. Quality

```{r echo=FALSE, warning=FALSE}
boxplot_and_median("chlorides")

boxplot_and_median("chlorides")+
  ylim(NA, quantile(wines$chlorides, 0.90))

```

A slight relation. Less chlorides means higher quality.

## Free sulfur dioxide vs. Quality

```{r echo=FALSE}
boxplot_and_median("free.sulfur.dioxide")

```

The middle quality classes seem to have higher free sulfur dioxide than both the low and high quality.

According to the dataset description, when free SO2 is lower than 50 ppm (~ 50 mg/L), it is undetectable. We can see in the following plot that very few wines are above this threshold which leads us to think that the variations seen in this plot are not related to an effect of the free SO2, but to the non balanced distribution of wines across the quality classes.

```{r echo=FALSE}
make_box_with_dots("free.sulfur.dioxide")+
  geom_hline(yintercept = 50, color = "#F1C40F", linetype = 2, size = 1.5)

```



## Total sulfur dioxide  vs. Quality

```{r echo = FALSE, warning=FALSE}
boxplot_and_median("total.sulfur.dioxide")+
  ylim(NA, 200)
```

Similar relation as with free sulfur dioxide. The middle classes have higher concentration than both the low and high.

## Density vs. Quality

```{r echo=FALSE}
boxplot_and_median("density")

```

Lower density means higher quality. From the dataset descriptions we know that the density will depend on the percentage of alcohol and sugar content. We should check those relationships later.


## pH vs. Quality

```{r echo=FALSE}
boxplot_and_median("pH")

```

There seems to be a trend of higher quality with lower pH. Higher quality with more acid content? We should check correlations between pH and the acidity levels.

## Sulphates vs. Quality

```{r echo=FALSE, warning=FALSE}
boxplot_and_median("sulphates")

boxplot_and_median("sulphates")+
  ylim(NA, quantile(wines$sulphates, 0.90))

```

Higher sulphates concentration means higher quality.

## Alcohol vs. Quality

```{r echo=FALSE}
boxplot_and_median("alcohol")

```

Besides the small downward bump in the quality class 5, the higher the alcohol content, the higher rated the wines get.

## Acidity and pH

```{r echo=FALSE}
ggplot(wines, aes(x = pH, y = fixed.acidity))+
  geom_point(alpha = 1/3)+
  scale_y_log10()

```

As expected the pH increases with the lower amount of acids. Fixed acidity accounts for most acids present in the wine.

```{r echo= FALSE}
ggplot(wines, aes(x = pH, y = citric.acid))+
  geom_point(alpha = 1/3)+
  scale_y_log10()

```

A similar relation is seen with the citric acid variable. But since the citric acid is at lower concentrations, the relation is not so strong. pH will be dominated by the other acids.

```{r echo=FALSE}
ggplot(wines, aes(x = pH, y = volatile.acidity))+
  geom_point(alpha = 1/3)+
  scale_y_log10()

```

The volatile acidity seems to have either no relation with the pH or a slight positive correlation.

Correlation coefficient:

```{r echo=FALSE}
cor.test(~ pH + log10(volatile.acidity), data = wines)

```

The correlation coefficient shows a weak positive correlation of volatile.acidity with the pH. Maybe when the volatile acids are present in higher concentration, the concentration of the remaining acids is lower and that contributes to the increase of pH.

```{r echo=FALSE}
ggplot(wines, aes(x = volatile.acidity, y = fixed.acidity))+
  geom_point(alpha = 1/2)

cor.test(~ fixed.acidity + volatile.acidity, data = wines)

```

We can see a weak negative correlation. On the plot, both variables seems to be have a natural limit on the lower sides. We have seen on the univariate plots that both are right skewed.

## Density, Sugar and Alcohol Content

The density of wine should be close to the water density, and will change depending on the percent of alcohol and sugar content.

```{r echo=FALSE, warning=FALSE}
ggplot(wines, aes(x = residual.sugar, y = density))+
  geom_point(alpha = 1/3)

ggplot(wines, aes(x = residual.sugar, y = density))+
  geom_point(alpha = 1/3)+
  xlim(NA, 4)
```

We see a increase of density with increase of residual sugar.

```{r echo=FALSE}
ggplot(wines, aes(x = alcohol, y = density))+
  geom_point(alpha = 1/3)

```

And we see a decrease of density with increase of alcohol content.

```{r echo=FALSE, warning=FALSE}
ggplot(wines, aes(x = alcohol, y = residual.sugar))+
  geom_point(alpha = 1/3)+geom_smooth()

ggplot(wines, aes(x = alcohol, y = residual.sugar))+
  geom_point(alpha = 1/3)+
  ylim(NA, 4)


cor.test(~ residual.sugar + alcohol, data = wines)
```

I was expecting a stronger correlation between the alcohol content and the residual sugar, since the alcohol comes from the fermentation of the sugars.

Maybe some of the wines are fortified with extra alcohol added that does not come from the fermentation of the sugar, or the yeast strains have different metabolic behaviors which do not allow to establish a linear relationship between sugar fermentation and alcohol production. Also, we don't know which grape types were used, which may have different sugar contents.


## Sulphates and sulfur oxide

Sulphate is an additive which can contribute to sulfur dioxide gas levels.

```{r echo=FALSE}
ggplot(wines, aes(x = sulphates, y = total.sulfur.dioxide))+
  geom_point(alpha = 1/5)

ggplot(wines, aes(x = sulphates, y = free.sulfur.dioxide))+
  geom_point(alpha = 1/5)

cor.test(~ total.sulfur.dioxide + sulphates, wines)
cor.test(~ free.sulfur.dioxide + sulphates, wines)

```

The relationship between sulphate levels and sulfur dioxide is very weak.



## Correlations with quality

```{r echo=FALSE}
cor(x = wines[1:11],
    y = as.numeric(wines$quality),
    method = "spearman")

```



# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

The wine quality is higher has stronger relationship with the volatile acidity, citric acid, sulphates and alcohol content. The correlation coefficients show us the strength of the relationship with the remaining variables.

```{r echo=FALSE}
cor(x = wines[1:11],
    y = as.numeric(wines$quality),
    method = "spearman")

```

For the free and total sulfur dioxide we have seen in the plots that the medium quality levels (5 and 6) have both higher content than the low and higher quality levels. This may hint at some interaction with the other variables.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

I observed the expected relation between the pH and acidity level.

It was interesting to observe the relation between the density and the alcohol and sugar content.

I was surprised by not finding a stronger relation between the residual sugar and alcohol level, since the alcohol comes from the fermentation of sugars.

### What was the strongest relationship you found?

The correlation coefficients show that the variable with the strongest relationship with quality is the alcohol content.




# MULTIVARIATE PLOTS SECTION
As a starting point let us have a look at a correlation matrix:

```{r echo=FALSE}
# http://www.r-bloggers.com/using-r-correlation-heatmap-with-ggplot2/



correlations <- as.data.frame(cor(data.frame(wines[1:11],
                                  quality_num = as.numeric(wines$quality)),
                                  method = "spearman")) %>%
                add_rownames(var = "Var1") %>% # need row names for the gathering
                gather(key = "Var2",
                      value = "correlation",
                      -Var1)

ggplot(data = correlations, aes(x = Var1, y = Var2))+
  geom_tile(aes(fill = correlation))+
  geom_text(aes(label = round(correlation, 3)), size = 2)+
  scale_fill_gradient2(low = "#E74C3C", mid = "white", high="#2471A3")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust = 1))


```

## Alcohol, volatile acidity and quality

Quality strongly correlates with alcohol and volatile.acidity. Volatile acidity comes from acetic acid which can give an unpleasant taste to the wine.

```{r echo=FALSE}
scatter_with_quality_color <- function(x, y){
  ggplot(wines, aes_q(x = as.name(x), y = as.name(y), color = ~ quality))+
    geom_point(alpha = 0.8, size = 2)+
    scale_color_brewer(type = "div", palette = "RdYlGn")
}

scatter_with_quality_color("alcohol", "volatile.acidity")

```

We can see the worse quality wines at low alcohol and high volatile acidity. The middle quality wines (5 and 6) can be found spread around a bit everywhere.

## Acidity, pH, quality

```{r echo=FALSE}
scatter_with_quality_color("fixed.acidity", "pH")+
  scale_x_log10()

```

There seems to be no pattern in the quality distribution here.

## Citric acid, alcohol, quality

```{r echo=FALSE}
scatter_with_quality_color("alcohol", "citric.acid")


```

The increase of both citric acid and alcohol tends to give higher quality wines. Nevertheless we see wines with quality 5 on a wide range of citric acid levels at low alcohol content and we see also high quality wines with low citric acid content.


## Alcohol and Sulphates

```{r echo=FALSE, warning=FALSE}
scatter_with_quality_color(x = "sulphates", y = "alcohol")

scatter_with_quality_color(x = "sulphates", y = "alcohol")+
  xlim(NA, 1.0)

```

For the range of sulphates between 0 and 1, alcohol and sulphates appear to have a positive correlation and higher alcohol combined with higher sulphates yields higher quality wines.

```{r echo=FALSE}
cor.test(~ alcohol + sulphates, data = wines)

```
# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

The main relationships explored were between the biggest correlators with quality.

We have seen how alcohol and volatile acidity relate with quality. Higher alcohol and lower acidity give in general better quality wines.

Higher amounts of citric acid combined with higher alcohol content yield the best wines.

Also with sulphates we see the same trend of better quality when both the alcohol and sulphates become higher.



```{r echo=FALSE, Multivariate_Plots,fig.width=12, fig.height=12}
 
Winet<-Wine.orig
Winet$quality<-Wine$good.wine
ggpairs(Winet, 
        aes(alpha=0.6),
        upper = list(continuous = wrap("cor", size = 8)),
        diag = list(continuous = "barDiag"),
        lower = list(continuous = "smooth"))
```




```{r echo=FALSE, Multivariate_Plots,fig.width=12, fig.height=12}
Winet$quality = as.ordered(Winet$quality)
ggpairs(Winet, aes(colour = quality, alpha = 0.4))
```

# Multivariate Analysis

We can observe the following, for each feature:

**fixed acidity** - besides poor quality, mean value and variance increases with quality;

**volatile acidity** - smaller means and smaller variance results in increasing quality;

**citric acid** - quality increases with the mean value;

**residual sugar** - highest quality has small mean, variance and less outliers;

**chlorides** - highest quality has smaller mean, variance and less outliers;

**free sulfur dioxide** - smaller mean and variance are for both small (3) and high quality (8);

**total sulfur dioxide** - smaller mean and variance are for both small (3) and high quality (8);

**density** - smaller mean, larger variance for higher quality;

**pH** - smaller values for higher quality;

**sulphates** - higher mean, smaller variance, less outliersfor higher quality;

**alcohol** - higher mean values, larger variance, less outliers for higher quality;


------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}

Wine$good.wine<-ifelse(Wine$quality>5.5,1,0)

ggplot(Wine,aes(x=good.wine,fill=factor(good.wine)))+
  geom_bar(stat = "count",position = "dodge")+
  scale_x_continuous(breaks = seq(0,1,1))+
  ggtitle("Distribution of Good/Bad Red Wines")+
  theme_classic()
```

Above plot shows what we have inferred previously, that good wines were not outnumbered by bad wines by a large margin. Most wines were mediocre (rated 5 or 6), but we could also see that there are some poor wines (3 or 4). A vast majority of good wines has a quality rating of 7.


### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

Wine %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value,fill=key)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins=sqrt(nrow(Wine))) +
    theme(legend.position="none") +
    scale_x_continuous(trans='log10')
```

### Description Two

In classification model, skewed distributions are undesirable, one method is to log transform the variable. We can show this effect by making the x axis on a log scale. The same plots are given above, but with the x axis on a log scale.

### Plot Three
```{r echo=FALSE, Plot_Three}
Wine.orig %>% cor() %>% corrplot.mixed(upper = "ellipse", tl.cex=.8,
                                       tl.pos = 'lt', number.cex = .8)
```

### Description Three

In feature selection, one method that we use is the correlation matrix and we choose the variables that are highly correlated with our interest variable

# Reflection

To wrap it all up, this has been a good investigation of red wine quality. Further analysis can still be made with the data and classification models can also be built (eg. Logistic model, decision trees or SVM). But for this project, we only performed a DEA which the convenient step to perform before building any statistical analysis.
